---
title: "Assignment 05"
author: "Maisy Christians"
date: "`r Sys.Date()`"
format: 
    pdf: default
bibliography: ref.bib
---

```{r}
#| echo: false
#| message: false

library(tidyverse)
```

```{r}
#| echo: false
#| message: false

wdi <- read.csv("wdi.csv")

#head(wdi)
```


## Indicator 1

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Source: World Development Indicators (2022)"

ggplot(wdi, aes(x = log10(gdp_per_capita), y = measles_immunisation_rate)) +
    geom_point(size = 3, alpha = 0.5) +
    theme_minimal() +
    labs(title = "Measles immunization vs. GDP", x = "log(GDP per capita)", y = "Measles immunization rate")
```

There are no rich countries who also have low measles immunization.



## Indicator 2

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Source: World Development Indicators (2022)"

ggplot(wdi, aes(x = health_expenditure_gdp_share, y = education_expenditure_gdp_share)) +
    geom_point(size = 3, alpha = 0.5) +
    theme_minimal() +
    labs(title = "Education expenditure vs. Health expenditure", x = "Education expenditure (GDP %)", y = "Health expenditure (GDP %)")
```

Educational expenditure and health expenditure have a direct relationship [@muarginean2014public].


## Indicator 3

```{r}
#| echo: false
#| warning: false
#| fig-cap: "Source: World Development Indicators (2022)"

wdi_3 <- wdi %>%
mutate(life_exp_group = case_when(
    life_expectancy < 70 ~ "<70",
    (life_expectancy >= 70 & life_expectancy <= 80) ~ "70-80",
    life_expectancy > 80 ~ ">80")) %>%
group_by(life_exp_group) %>%
summarize(avg = mean(gdp_per_capita, na.rm = TRUE), count = n())

wdi_3$life_exp_group <- factor(wdi_3$life_exp_group,
    levels = c("<70", "70-80", ">80"))

ggplot(wdi_3, aes(x = life_exp_group, y = avg)) +
    geom_col(aes(fill = life_exp_group), show.legend = FALSE) +
    theme_minimal() +
    labs(title = "GDP vs. Life expectancy", x = "Life expectancy (years)", y = "Average GDP per capita")
```

GDP is higher for countries with greater life expectancies [@kabir2008determinants].
See @tbl-1 for numbers used.


## Table with key statistics

|Life expectancy (years) | # of countries| Average GDP per capita |
|:---|:-----:|:-------:|
|<70| 65|$2,955|
|70-80|105|$13,562|
|>80|47|$61,484|
: {#tbl-1}


## Bibliography